---
title: "Membench"
author: "K. Siek"
date: "November 23, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Membench

A micro benchmark to show the memory usage of UFOs. This notebook runs and graphs memory consumption.

### Compiling and running the benchmark

Compile the benchmark.

```{bash}
make
```

Run the benchmark with parameters:

  - `size`  Size of the UFO (in elements of type `int64_t`),
  - `writes` The number of read operations between writes, zero for read-only,
  - `min-load` UFO chunk size (in bytes),
  - `high-water-mark`The threshold for memory used by materialized chunks before GC starts (in bytes),
  - `low-water-mark` Once GC starts, it removes chunks until this much memory is left (in bytes),

``` {bash}
./membench --size=$((10*1000*1000)) --writes=0 --min-load=$((1024 * 1024 / 8)) --high-water-mark=$((10 * 1024 * 1024)) --low-water-mark=$((1 * 1024 * 1024))
```

### Memory and disk usage graph

Load R libraries:
```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(gdata)
library(scales)
```
```{r}
KB <- function(x) { x * 1024 }
MB <- function(x) { x * 1024 * 1024 }
GB <- function(x) { x * 1024 * 1024 * 1024 }

# labels_nanoseconds <- function(x) {
#   ifelse(abs(x) < 1000, paste0(x, "ns"),
#          ifelse(abs(x) < 1000 * 1000, paste0(x, "µs"), 
#                 ifelse(x < 1000 * 1000, paste0(x, "ms"))
#                 )
#   )
# }
```

Ingest the data
```{r}
data <- read_csv("membench.csv")
```

```{r fig.width=20, fig.height=10}
ggplot(data, aes(x=timestamp)) + 
  geom_area(aes(y=memory_usage), fill="#69b3a2", alpha=0.4) +
  geom_line(aes(y=intended_memory_usage), color="#69b3a2", size=2) +
  theme_linedraw() + 
  scale_y_continuous(breaks=MB(10*(0:8)), labels = function(x) humanReadable(x, standard = "IEC", sep = "")) +
  scale_x_continuous(labels = function(x) humanReadable(x, standard = "SI", sep = "", units="")) +
  geom_line(aes(y=disk_usage)) +
  theme(axis.title.y = element_blank())
  
  #geom_line(aes(y=memory_usage)) #+ geom_line(aes(y=disk_usage)) #+ geom_line(aes(y=intended_memory_usage)) + geom_line(aes(y=apparent_memory_usage))
```