---
title: "Membench"
author: "K. Siek"
date: "November 23, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Membench

A micro benchmark to show the memory usage of UFOs. This notebook runs and graphs memory consumption.

### Compiling and running the benchmark

Compile the benchmark.

```{bash}
make
```

Run the benchmark with parameters:

  - `ufos` Number of vectors (1 or 2)
  - `size`  Size of the UFO (in elements of type `int64_t`),
  - `writes` The number of read operations between writes, zero for read-only,
  - `min-load` UFO chunk size (in bytes),
  - `high-water-mark`The threshold for memory used by materialized chunks before GC starts (in bytes),
  - `low-water-mark` Once GC starts, it removes chunks until this much memory is left (in bytes),

``` {bash}
./membench --size=$((10*1000*1000)) --writes=100 --min-load=$((10*4096)) --high-water-mark=$((20 * 1024 * 1024)) --low-water-mark=$((10 * 1024 * 1024)) --ufos=2
```

### Memory and disk usage graph

Load R libraries:
```{r}
library(dplyr)
library(readr)
library(ggplot2)
library(gdata)
library(scales)
library(cowplot)

library(extrafont)
loadfonts()
```

Helper functions
```{r}
KB <- function(x) { x * 1024 }
MB <- function(x) { x * 1024 * 1024 }
GB <- function(x) { x * 1024 * 1024 * 1024 }

find_memory_order <- function(x) {
  ifelse(x >= GB(10), GB, ifelse(x >= MB(10), MB, ifelse(x >= KB(10), KB, function(x) x)))
}

# labels_nanoseconds <- function(x) {
#   ifelse(abs(x) < 1000, paste0(x, "ns"),
#          ifelse(abs(x) < 1000 * 1000, paste0(x, "µs"), 
#                 ifelse(x < 1000 * 1000, paste0(x, "ms"))
#                 )
#   )
# }

drop_zeros <- function(x) ifelse(x != 0, x, NA)
```

Ingest the data
```{r}
data <- read_csv("membench.csv")
parameters <- read_csv("parameters.csv")
```

```{r fig.width=20, fig.height=10}
memory_breaks <- function(step=20) {
  max_memory_usage <- max(data$memory_usage, data$intended_memory_usage, data$disk_usage, data$apparent_memory_usage, parameters$high_water_mark)
  memory_order <- find_memory_order(max_memory_usage)
  memory_order((0:(max_memory_usage / memory_order(1) / step)) * step)
}

memory_usage_color <- "#457b9d"
ufo_1_memory_usage_color <- "#f1faee"
ufo_0_memory_usage_color <- "#a8dadc"
disk_usage_color <- "#f4a261"
ufo_perceived_color <- "black"
ufo_actual_color <- "black"
high_watermark_color <- "#e63946"
low_watermark_color <- "#e63946"
ufo_0_created_color <- "grey"
ufo_1_created_color <- "grey"

graph_data <- data %>% mutate(memory_usage = drop_zeros(memory_usage),
                              disk_usage = drop_zeros(disk_usage),
                              ufo_0_memory_usage = drop_zeros(ufo_0_memory_usage),
                              ufo_1_memory_usage = drop_zeros(ufo_1_memory_usage))

offset <- 1000000

ufo_0_created_y <- unique(data$apparent_memory_usage[data$apparent_memory_usage != 0])[1]
ufo_1_created_y <- unique(data$apparent_memory_usage[data$apparent_memory_usage != 0])[2]
ufo_0_created_x <- data %>%filter(data$apparent_memory_usage == ufo_0_created_y) %>% pull(timestamp) %>% min 
ufo_1_created_x <- data %>%filter(data$apparent_memory_usage == ufo_1_created_y) %>% pull(timestamp) %>% min 

p <- ggplot(graph_data, aes(x=timestamp)) + 
  geom_rect(aes(xmin=timestamp,xmax=lead(timestamp),ymin=0,ymax=disk_usage), fill=disk_usage_color, color=disk_usage_color, alpha=1) +
  geom_rect(aes(xmin=timestamp,xmax=lead(timestamp),ymin=0,ymax=memory_usage), fill=memory_usage_color, color=memory_usage_color, alpha=1) +
  geom_step(aes(y=memory_usage), color=memory_usage_color) +
  geom_step(aes(y=ufo_0_memory_usage), color=ufo_0_memory_usage_color, size=0.75) + #linetype = "dotted",
  geom_step(aes(y=ufo_1_memory_usage), color=ufo_1_memory_usage_color, size=0.75) +
  geom_step(aes(y=disk_usage), color=disk_usage_color, size=1) +
  
  # UFO sizes
  geom_step(aes(y=apparent_memory_usage), color=ufo_actual_color, size=0.5, linetype = "dashed") +
  # Threshods
  geom_hline(yintercept=parameters$high_water_mark, color=high_watermark_color, linetype = "dashed") +
  geom_text(aes(y=parameters$high_water_mark, x=unique(min(timestamp))+offset, label="high watermark"), color=high_watermark_color, vjust=-0.75, hjust=0, check_overlap = TRUE) +
  geom_hline(yintercept=parameters$low_water_mark, color=low_watermark_color, linetype = "dashed") +
  geom_text(aes(y=parameters$low_water_mark, x=unique(min(timestamp))+offset, label="low watermark"), color=low_watermark_color, vjust=-0.75, hjust=0, check_overlap = TRUE) +

  geom_text(aes(y=ufo_0_created_y, x=ufo_0_created_x + offset, label=paste0("ufo::new(", label_number_si()(parameters$size / 1), " * int64)")), vjust=-0.75, hjust=0, check_overlap = TRUE) +
  geom_text(aes(y=ufo_1_created_y, x=ufo_1_created_x + offset, label=paste0("ufo::new(", label_number_si()(parameters$size / 2), " * int64)")), vjust=-0.75, hjust=0, check_overlap = TRUE) +

  geom_text(aes(y=max(unique(data$apparent_memory_usage)), x=max(unique(timestamp)) - offset, label="perceived memory usage"), vjust=-0.5, hjust=1, check_overlap = TRUE) +  
  geom_text(aes(y=max(unique(data$disk_usage)), x=max(unique(timestamp)) - offset, label=paste0("total disk usage"), fontface="bold"), color=disk_usage_color, vjust=-0.75, hjust=1, check_overlap = TRUE) +
  geom_text(aes(y=max(unique(data$memory_usage)), x=max(unique(timestamp)) - offset, label=paste0("total memory usage"), fontface="bold"), color=memory_usage_color, vjust=-0.75, hjust=1, check_overlap = TRUE) +
  
  #geom_text(aes(y=0, x=data %>% filter(ufo_0_memory_usage > 0) %>% pull(timestamp) %>% min + 2 * offset, label="ufo 1 memory usage"), color=ufo_0_memory_usage_color, vjust=-0.5, hjust=0, check_overlap = TRUE) +
  #geom_text(aes(y=0, x=data %>% filter(ufo_1_memory_usage > 0) %>% pull(timestamp) %>% min + 2 * offset, label="ufo 2 memory usage"), color=ufo_1_memory_usage_color, vjust=-0.5, hjust=0, check_overlap = TRUE) +
  
  geom_text(aes(y=0, x=max(unique(timestamp)) - offset, label="ufo 1 memory usage"), color=ufo_0_memory_usage_color,  vjust=-2.5, hjust=1, check_overlap = TRUE) +  
  geom_text(aes(y=0, x=max(unique(timestamp)) - offset, label="ufo 2 memory usage"), color=ufo_1_memory_usage_color,  vjust=-0.75, hjust=1, check_overlap = TRUE) +  
  
  # Theming
  #guides(color = guide_legend(title="XXX")) +
  theme_linedraw() + 
  scale_y_continuous(labels = function(x) humanReadable(x, standard = "IEC", sep = ""),  expand = expansion(mult = c(0, .05))) + # breaks=memory_breaks(20), 
  scale_x_continuous(labels = function(x) paste0(floor(x / 1000 / 1000), "ms"), expand = expansion(mult = c(.005, 0))) +
  theme(axis.title.y = element_blank()) +
  scale_color_manual(values = colors) +
  xlab("elapsed time")

ggsave(plot=p, filename="membench.pdf", width=20, height=10, dpi = "print")
ggsave(plot=p, filename="membench.png", width=20, height=10, dpi = "print")
ggsave(plot=p, filename="membench.svg", device="svg", width=20, height=10, dpi = "print")

p
```
